---
title: "Shiny Training Cheat Sheet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```

## Create a new Shiny app

File > New File > Shiny Web App...
Enter name for app, set as single file app (if not already selected) and directory.
Finally, click Create.

## Input elements

In `{shiny}` input elements follow a pattern like below, where the ellipsis `...` means any other arguments, and these arguments will depend on the type of input.

`{typeOfInput}Input(inputId, label, ...)`

For example

`textInput("nameOfInput", "Label for input")`

Any element you create in `{shiny}` is translated to HTML.

```{r input_becomes_html}
print(textInput("nameOfInput", "Label for input"))
```


### Adding an input

To add an input element, you just need to call the desired input function in your `ui` function, supplying as a minimum the `inputId` and `label` arguments. If you want to check what arguments are available for a specific input, remember to use the `?` in front of a function or press `F1` with the input function selected in your code.

A simple example is

```
ui <- fluidPage(
  textInput("myInput", "I'm a text input!")
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Output elements

Output elements in `{shiny}` follow a pattern like

`{typeOfOutput}Output(outputId, ...)`

Just like inputs, the additional arguments vary with the output type.

For example

`textOutput("nameOfOutput")`

As mentioned for inputs, an output element becomes HTML when the app is run.

```{r output_becomes_html}
print(textOutput("nameOfOutput"))
```

### Accessing an input

To access, or refer to, an input, we normally use the syntax

`input$nameOfInput`

Shiny adds all inputs you declare to an internal named list called `input`. Because it is a list, you can also access inputs like

`input[["nameOfInput"]]`

### Adding an output

Just as inputs are added to the `ui` function, so too with outputs. Continuing the simple app already shown, we could add an output like below.

```
ui <- fluidPage(
  textInput("myInput", "I'm a text input!"),
  textOutput("myOutput")
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

However, running this app would look just the same whether the output is there or not. This is because we have not told `{shiny}` just _what_ the output should display.

### Accessing an output

As you might expect, outputs are handled similarly to inputs; you access them like

`output$nameOfOutput`

or

`output[["nameOfOutput"]]`

### Adding content to an output

For an output to actually display anything, you must add some content to it. This is done by using one of the family of `render` functions. They follow a pattern like

`render{typeOfElement}(expr, ...)`

For example

`renderText("I'll appear on the page when the app is run")`

To assign this to an output we use syntax like

`output$myOutput <- renderText({"Some output"})`

These output assignments belong in the `server` function of the app, like below

```
ui <- fluidPage(
  textInput("myInput", "I'm a text input!"),
  textOutput("myOutput")
)

server <- function(input, output, session) {
  output$myOutput <- renderText({"...and I'm a text output!"})
}

shinyApp(ui, server)
```

## Reactivity

In `{shiny}` we have a powerful technique for writing reactive apps. Using reactive expressions we can create expressions whose output depends on their input. This allows for dynamic behaviour, as a user makes choices with the inputs of the app.

To create a reactive expression use the syntax

```
reactiveData <- reactive({

  return_data %>%
  ...
})
```

To use this expression, we use the same syntax as we do for calling a function with no arguments, `reactiveData()`

For example

```
output$table <- renderDT({

  returnData() %>%
  ...
})
```

Note that reactive expressions can only be used in a __reactive context__. This includes the `render` family of functions.

Continuing the simple app, we could have

```
ui <- fluidPage(
  textInput("myInput", "I'm a text input!"),
  textOutput("myOutput")
)

server <- function(input, output, session) {
  reactiveText <- reactive({
    upper(input$myInput)
  })
  output$myOutput <- renderText({
    reactiveText() # Reference a reactive expression - it is not a function!
  })
}

shinyApp(ui, server)
```

## Debugging

As Hadley Wickham says many times in his numerous books and blogs

> Debugging Shiny code is hard

This is often due to the complexities introduced by reactivity.

Some tips for debugging include

- Start simple
- Make sure your code works outside of Shiny! One way of writing Shiny apps is known as **markdown first**; you write your code in RMarkdown initially, and only once everything (or some logical unit of the whole code) is working, only then create (or add to) a `{shiny}` app
- Use the `cat()` or `print()` functions
- Breakpoints and `browser()`
- Run `options(shiny.reactlog=TRUE)`, run the app and press `Ctrl+F3` for an interactive summary of reactive activity
