---
title: "Shiny training- session 2"
author: "Chris Beeley"
date: "November 2020"
output: 
  ioslides_presentation:
    css: temp.css
runtime: shiny
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Introduction

* Recap
* *Application 3- A and E*
* Tabset panels
* Conditional UI
* Render UI
* Validate() and req()

## Recap

* Application setup
* Reactivity
* Reactive expressions
* Optional:
    * NSE
    * Bookmarks

## Tabset panels

* Tabset panels are useful if you want multiple outputs but not all on the same page
* If you give them an id you can test for which is selected too which is very useful.

```{r, eval = FALSE}

mainPanel(
  tabsetPanel(id = "tabset",
              tabPanel("Label", outputFunction("nameOfOutput")),
              tabPanel("Label2", outputFunction2("nameOfOutput2"))
  )
)

```

## Let's build an app!

* Open the a_and_e_code_first.R file
* Contains a template for the tabPanel
* Contains all server code
* You need to lay out and app and add:
* dateRangeInput (to select date)
* selectInput (to select Trust)
* plotOutput (to place graph in one tab)
* leafletOutput (to place map in the other tab)
* I suggest you use the wizard to produce a new app
* File... New... Shiny web app... Application name... Single... Choose directory
* Note that you need to select a Trust otherwise you will get an error (we're going to fix this later)
* Answer is in a_and_e_first folder. Use sparingly!

## Add more features

* Conditional UI
* Dynamic UI
* Validating the state of the application

## Dynamic UI

* Producing dynamic UI is useful if you have a very large set of inputs but only want some depending on other inputs
* In the application we're building today it's useful because not every org has data every year

## Conditional panel

* Conditional panels are useful if you only want to show your UI element when a certain condition pertains
* Note that you provide JavaScript code in a string
* Easier than it sounds
* input$myInput becomes input.myInput

```{r, eval = FALSE}

# ui

conditionalPanel(
  condition = "input.myInput == 'yes'",
      selectInput("trust",
                "Select Trust",
                choices = unique(ae_attendances$Name),
                multiple = TRUE)
)

```

```{r, eval = FALSE}

# ui

uiOutput("trustControl")

# server

output$trustControl <- renderUI({
  
  selectInput(
    ###
  )
})

```

## Validating state of application

* Remember the nasty red error message we got before we selected a trust?
* Shiny has two ways of avoiding these
* req()
* req() will refrain from running a block of code until whatever is inside it takes a valid value
* validate() will also refrain from running a block of code but will also write a nice error message

```{r eval = FALSE}

# req()

req(input$trust) # this will not run until input$trust has a value

# validate()

validate(
  need(input$trust, "Please select a trust"),
  need(input$somethingElse "Please select a value for this input too")
)

```

## Let's add to the application

* Open the a_and_e_code_second.R file
* Conditional UI
* Dynamic UI
* Validating the state of the application
* The answer is in the a_and_e_second folder. Use sparingly!

## Layout

* fluidRow()
* column(width = 6)
* Columns add up to 12
* That's it!

## Example

```{r}

ui <- fluidPage(
  fluidRow(
    column(width = 6, wellPanel(p("This appears in one column"))
    ),
    column(width = 6, wellPanel(p("Then this in another column"))
    )
  ),
  fluidRow(
    column(width = 3, wellPanel(p("Then a thin column underneath the top row"))),
    column(width = 9, wellPanel(p("Then a wide column next to that one")))
  )
)

server <- function(input, output) {}

shinyApp(ui = ui, server = server)

```

## Let's add to the application

* Open the a_and_e_code-third.R file
* Have a go at laying out the application yourself
* Answer is a_and_e_code_third

## Bonus activity

* Use renderUI() to make the dateRangeInput() start at the beginning of the data and finish at the end
* In this application obviously the data is not dynamic so it isn't necessary
* But a real application might very well include this functionality
* Or explore the other layout functions from the previous exercise 
    * splitLayout()
    * verticalLayout()
    * navbarPage()
    * navlistPanel()