---
title: "Shiny training- session 3"
author: "Chris Beeley"
date: "November 2020"
output: 
  ioslides_presentation:
    css: temp.css
runtime: shiny
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

## Introduction 

Application 3- enhance the A + E one

* RMarkdown
* shinydashboard
* Interacting with a Shiny plot (TBH this is kind of hard)

## RMarkdown

```{r, eval = FALSE}

# UI

downloadButton("downloadReport", "Download report")

# server

output$downloadReport <- downloadHandler(
  filename = "report.docx",
  content = function(file){
    
    params <- list(date_from = input$date[1],
                   date_to = input$date[2],
                   trust = input$trust)
    
    render(paste0("report.Rmd"), output_format = "word_document",
           output_file = "report.docx", quiet = TRUE, params = params,
           envir = new.env(parent = globalenv()))
    
    # copy docx to 'file'
    file.copy(paste0("report.docx"), file, overwrite = TRUE)
  }
)

```

## In the report

```{r, eval = FALSE}

# YAML

params: 
  date_from : NA
  date_to : NA
  trust : NA
output: html_document

# R

load("ae_attendances.RData")

report_data <- ae_attendances %>% 
  filter(period >= date_from, period <= date_to,
         Name == trust)

```

* If they don't select a Trust first they will get an error
* Multiple ways of doing this
* Have a look in the answer sheet to see one after you've had a go yourself
* Helper code in a_and_e_code_fourth.R
* The answer is in the a_and_e_fourth folder. Use sparingly!

## Shiny dashboard

```{r, eval = FALSE}

# this runs and produces a totally blank dashboard

library(shiny)
library(shinydashboard)

ui <- dashboardPage(
  dashboardHeader(),
  dashboardSidebar(),
  dashboardBody()
)

server <- function(input, output) { }

shinyApp(ui, server)

```

## Value boxes

```{r, eval = FALSE}

# https://rstudio.github.io/shinydashboard/structure.html#valuebox

# ui

valueBox()

# Dynamic valueBoxes

valueBoxOutput("daysBox")

# server

output$daysBox <- renderValueBox({
  valueBox(
  )
})

```

* Your turn!
* Open a_and_e_code_fifth.R
* [https://rstudio.github.io/shinydashboard/index.html](https://rstudio.github.io/shinydashboard/index.html)
* Answers in a_and_e_dashboard. Use sparingly

## Finally

* Either make both plots interactive (hard- answers in a_and_e_interactive)
* Make dynamic value boxes based on interacting with a plot (extreme- no answers)
* You may wish to take one or all of these away as homework
